What if you drop that? Modeling children's curiosity about physical interactions

### 2020-01-29 paper/paper.Rmd
Update paper.Rmd

The goal of the current paper is to help close this gap by proposing a framework for advancing our theoretical understanding of curiosity that combines three key ingredients: _first_, we present a computational theory of how a lerner's knowledge state interacts with ongoing perceptual experiences to give rise to curiosity-driven exploratory behavior; _second_, we present an empirical investigation of the pattern of actions taken by children in a novel physical exploration task; and _third_, we evaluate the quantitative correspondence between the exploratory actions taken by an artificial agent that instantiates our computational theory and those taken by human children on the same set of tasks.

@@

### Tower Task

<!-- 1. height x continuous age -->
<!-- First we examine tower height as a function of age. Y/N we find tower height increases with age. -->
```{r}
# cor.test(hum$height_tallest, hum$age_rounded)
```

```{r, include=F} 
# ch_height <- hum %>% group_by(Age_Group, height_tallest) %>% 
#   summarise(n = n()) %>%
#   mutate(freq = n / sum(n)) 
# 
# ggplot(ch_height, aes(x=age_group, y=height_tallest)) + 
#   geom_smooth(method="lm", level=0.90) + geom_point(alpha=.8, aes(size=n)) +
#   scale_shape_manual(values=10:19) +
#   xlab("Age") + ylab("Height_Tallest") + ylim(0,10) +
#   langcog::scale_fill_solarized() + ggthemes::theme_few()
```

@@

<!-- 3. The structural components of the towers -->
<!-- -support x supported -->
<!-- -drop object probabilities correlate with objects used in tower? -->
<!-- -first base selection as proxy of most interesting object  -->
<!-- - base and top distribution  -->

>>>>

The goal of the current paper is to help close this gap by proposing a framework for advancing our theoretical understanding of curiosity that combines three key ingredients: _first_, we present a computational theory of how a learner's knowledge state interacts with ongoing perceptual experiences to give rise to curiosity-driven exploratory behavior; _second_, we present an empirical investigation of the pattern of actions taken by children in a novel physical exploration task; and _third_, we evaluate the quantitative correspondence between the exploratory actions taken by an artificial agent that instantiates our computational theory and those taken by human children on the same set of tasks.

@@

### Tower Task Results

```{r}
require(here)
require(tidyboot)
library(ggplot2)
require(tidyverse)

tower = read.csv2("../analysis/human-data/curiobaby_drop-data - cool tower.csv", sep=',', header=T, stringsAsFactors=F)
hum = subset(tower, exclude!="Y")

length(unique(tower$SID)) #61
length(unique(hum$SID)) # 51 subjects

hum$age_group = as.numeric(hum$age_group)
hum$age_rounded = as.numeric(hum$age_rounded)
hum$height_tallest = as.numeric(hum$height_tallest)
hum$tallest_objects = as.character(hum$tallest_objects)
hum$base_tallest = as.character(hum$base_tallest)

1. height x continuous age
First we examine tower height as a function of age. Yes we find tower height increases with age.
```{r}
cor.test(hum$height_tallest, hum$age_rounded)
```

```{r}
ch_height <- hum %>% group_by(age_rounded, height_tallest) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

ggplot(ch_height, aes(x=age_rounded, y=height_tallest)) +
  geom_smooth(method="lm", level=0.90) + geom_point(alpha=.8) +
  scale_shape_manual(values=10:19) +
  xlab("Age") + ylab("Height_Tallest") + ylim(0,10) +
  langcog::scale_fill_solarized() + ggthemes::theme_few()
```

2. support x supported (counts)
```{r}
tower = subset(tower, tallest_objects!="" & tallest_objects!="pentagon+octahedron, trig prism") # 5 kids who arranged horizontally and 1 kid chose 2 - exclude
df = data.frame(support=NA, supported=NA, n=NA)
for(s in 1:nrow(tower)) {
  objs = unlist(str_split(tower[s,]$tallest_objects, ", "))
  print(objs)
  for(i in 2:length(objs)) {
    df = rbind(df, cbind(support=objs[i-1], supported=objs[i], n=1))
  }
}
df = na.omit(df)
hm <- df %>% group_by(support, supported) %>% count(N=n()) 
hm %>% ggplot(aes(x=support, y=supported)) + geom_tile(aes(fill=N)) + geom_text(aes(label = N)) +
  scale_fill_gradient(low = "white", high = "steelblue") + ggthemes::theme_few()
```

3. support x supported (probabilities)
pr (supported object | support object)
```{r}
tower = subset(tower, tallest_objects!="" & tallest_objects!="pentagon+octahedron, trig prism") # 5 kids who arranged horizontally and 1 kid chose 2 - exclude
df = data.frame(support=NA, supported=NA, n=NA)
for(s in 1:nrow(tower)) {
  objs = unlist(str_split(tower[s,]$tallest_objects, ", "))
  print(objs)
  for(i in 2:length(objs)) {
    df = rbind(df, cbind(support=objs[i-1], supported=objs[i], n=1))
  }
}
df = na.omit(df)
hm1 <- df %>% 
  group_by(support, supported) %>% 
  summarise(n = n()) %>% 
  group_by(support) %>%
  mutate(freq = n / sum(n))
hm1 %>% ggplot(aes(x=support, y=supported)) + geom_tile(aes(fill=freq)) + geom_text(aes(label = round(freq, 2))) +
  scale_fill_gradient(low = "white", high = "steelblue") + ggthemes::theme_few()
```

4. base and top distribution
```{r}
ch_extremes <- hum %>% 
  group_by(base_tallest) %>% 
  count(N=n()) %>%
  filter(base_tallest!="" & base_tallest!="pentagon+octahedron")

ch_extremes %>% 
  ggplot(aes(x = base_tallest, y = N)) + geom_bar(stat = "identity")

@@

<!-- 3. The structural components of the towers -->
<!-- -support x supported -->
<!-- -drop object probabilities correlate with objects used in tower? -->
<!-- -first base selection as proxy of most interesting object -->
<!-- -base and top distribution -->
